### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Initial Commit

---
title: "outlier_exploration"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(car)
```

```{r}
full_model <- lm(sleep_duration ~., sleep)
```

### Outliers

```{r}
plot(full_model)
```

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

```

### Multicollinearity

### 2025-04-26 data/data_wrangling.R

@@ -1,0 +1,25 @@
+library(readr)
+library(tidyverse)
+library(janitor)
+
+sleep <- read_csv("data/Sleep_health_and_lifestyle_dataset.csv") %>% 
+  clean_names()
+
+glimpse(sleep)
+
+sleep <- sleep %>%
+  separate(blood_pressure, into = c("systolic", "diastolic"), sep = "/", convert = TRUE)
+
+sleep$stress_level <- as.factor(sleep$stress_level)
+
+sleep <- sleep %>%
+  mutate(
+    systolic = as.numeric(systolic),
+    diastolic = as.numeric(diastolic),
+    MAP = (2 * diastolic + systolic) / 3
+  )
+
+sleep <- sleep %>% 
+  select(!c(person_id, systolic, diastolic))
+
+glimpse(sleep)

### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Ethan: Removing potential outliers, R^2 value increased significantly

@@
 full_model <- lm(sleep_duration ~., sleep)
+summary(full_model)
 ```
@@
 
+# Remove the suspicious rows
+sleep_clean <- sleep[-suspicious_indices, ]
+
+# Refit the model excluding suspicious rows 
+full_model_clean <- lm(sleep_duration ~ ., data = sleep_clean)
+
+# Plot diagnostics again on the cleaned model
+par(mfrow = c(2, 2))
+plot(full_model_clean)
+
+# Comparison of model summaries
+summary(full_model)        # Old model (with potential outliers)  R^2  = 0.9592
+summary(full_model_clean)  # New model (without potential outliers)  R^2  = 0.9929
+
 ```

### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Ethan: Removed aliased variable (BMI), obtained VIF values to observe multicolinearity, used step-wise AIC variable selection which removed quality_of_sleep variable, obtained VIF values of new model and were improved dramatically.

@@
 
+```{r}
+
+```
+
 ### Multicollinearity
+
+```{r}
+library(car)
+
+# Calculate VIFs on the cleaned model
+vif_values <- vif(full_model_clean)
+
+# Running VIF gave error (aliased variables are present in the model), so find out which are aliased 
+alias(full_model_clean) # BMI->Overweight is perfectly aliased with other variables in the model
+
+# Refit the model without BMI
+full_model_clean2 <- lm(sleep_duration ~ gender + age + occupation + quality_of_sleep +
+                        physical_activity_level + stress_level + heart_rate +
+                        daily_steps + sleep_disorder + MAP, data = sleep_clean)
+
+# Calculate VIFs on the cleaned model, with BMI removed 
+vif_values <- vif(full_model_clean2)
+vif_values     # Significant multicolinearity present, variable selection is next. Start with step-wise AIC variable selection
+
+```
+
+### Variable Selection
+
+```{r}
+# Step-wise AIC Variable Selection
+
+library(MASS)
+
+step_model <- stepAIC(full_model_clean2, direction = "both", trace = FALSE)
+
+# Check final selected model
+summary(step_model)
+
+# Observing which variables were removed from full model 
+formula(full_model_clean2) 
+formula(step_model)      # Variable removed: quality_of_sleep
+
+# Check VIF of the step-wise model
+vif(step_model)      # VIF values have decreased significantly after step-wise variable selection
+
+```

### 2025-05-03 paper/paper.qmd
paper

---
title: "Multiple Linear Regression on Sleep Data"
author:
  - name: Ethan Newcomb
  - name: Kyubin Im
  - name: Matthew Murnane
format:
  pdf:
    mainfont: "Times New Roman"
indent: true
fontsize: "12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
#Packages
library(tidyverse)
library(dplyr)
library(ggthemes)
library(tidyr)
library(knitr)
library(kableExtra)
library(car)
library(MASS)
library(readr)
library(janitor)
```

```{r include=FALSE}
#Data Wrangling Code
sleep <- read_csv("../data/Sleep_health_and_lifestyle_dataset.csv") %>%  
  clean_names()

sleep <- sleep %>%
  separate(blood_pressure, into = c("systolic", "diastolic"), sep = "/", convert = TRUE)

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$quality_of_sleep <- as.factor(sleep$quality_of_sleep)

sleep <- sleep %>%
  mutate(systolic = as.numeric(systolic),
         diastolic = as.numeric(diastolic),
         PP = systolic-diastolic)

sleep <- sleep %>%
  dplyr::select(!c(person_id, systolic, diastolic))
```

\newpage

\noindent\textbf{\Large Motivation}

\vspace{0.5em}

Sleep is important and people are always wondering how they can improve their sleep. Everyone knows how annoying it is to be tossing and turning all night long. Lifestyle and general health seem to be important indicators. In this report we explore a sleep data set and see what lifestyle and health variables have a significant relationship with sleep duration.

\vspace{1.5em}

\noindent\textbf{\Large About the Data}

```{r}
#Table 1: Variables in Sleep
variable_names <- matrix(c(
  "Gender", "Age", "Occupation",
  "Sleep Duration (hours)", "Quality of Sleep (scale: 1–10)", "Physical Activity Level (minutes/day)",
  "Stress Level (scale: 1–10)", "BMI Category", "PP (Pulse Pressure)",
  "Heart Rate (bpm)", "Daily Steps", "Sleep Disorder"
), ncol = 3, byrow = TRUE)

kable(variable_names, col.names = NULL, caption = "Variables From Our Data")
```

Sleep Duration is a continuous variable and will be our response variable. The rest will be our covariates. Gender, Occupation, Quality of Sleep, Stress Level, BMI Category and Sleep Disorder are categorical variables. Physical Activity, MAP, and Daily steps are continuous variables. We have a total of 374 observations.

We had no missing data in the data set but we did have to make a mutation using `dplyr`. In the original data set there was a blood pressure variable that was encoded as a character but was a ratio, e.g. "$\frac{120}{80}$". We separated the blood pressure variable into two: Systolic and Diastolic. These two we suspected to be highly correlated but did not want to just drop one or the other. We had options of combining the two into something meaningful. One was Pulse Pressure (PP), which is calculated as $\text{PP}=\text{systolic}-\text{diastolic}$. PP is the force of the hearts contraction on the arteries. The other is called Mean Arterial Pressure (MAP), which is calculated as $\text{MAP}=\frac{\text{systolic}\cdot\text{diastolic}^2}{3}$. MAP is the average blood pressure throughout a cardiac cycle. We decided on PP because it was simpler to calculate and understand.

### 2025-05-03 paper/paper.tex
nothing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
    \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Multiple Linear Regression on Sleep Data},
  pdfauthor={Ethan Newcomb; Kyubin Im; Matthew Murnane},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Multiple Linear Regression on Sleep Data}
\author{Ethan Newcomb \and Kyubin Im \and Matthew Murnane}
\date{}

\begin{document}
\maketitle

\newpage

\noindent\textbf{\Large Motivation}

\vspace{0.5em}

Sleep is important and people are always wondering how they can improve
their sleep. Everyone knows how annoying it is to be tossing and turning
all night long. Lifestyle and general health seem to be important
indicators. In this report we explore a sleep data set and see what
lifestyle and health variables have a significant relationship with
sleep duration.

\vspace{1.5em}

\noindent\textbf{\Large About the Data}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2812}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3229}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3958}}@{}}
\caption{Variables From Our Data}\tabularnewline
\toprule\noalign{}
\endfirsthead
\endhead
\bottomrule\noalign{}
\endlastfoot
Gender & Age & Occupation \\
Sleep Duration (hours) & Quality of Sleep (scale: 1--10) & Physical
Activity Level (minutes/day) \\
Stress Level (scale: 1--10) & BMI Category & PP (Pulse Pressure) \\
Heart Rate (bpm) & Daily Steps & Sleep Disorder \\
\end{longtable}

Sleep Duration is a continuous variable and will be our response
variable. The rest will be our covariates. Gender, Occupation, Quality
of Sleep, Stress Level, BMI Category and Sleep Disorder are categorical
variables. Physical Activity, MAP, and Daily steps are continuous
variables. We have a total of 374 observations.

We had no missing data in the data set but we did have to make a
mutation using \texttt{dplyr}. In the original data set there was a
blood pressure variable that was encoded as a character but was a ratio,
e.g.~``\(\frac{120}{80}\)''. We separated the blood pressure variable
into two: Systolic and Diastolic. These two we suspected to be highly
correlated but did not want to just drop one or the other. We had
options of combining the two into something meaningful. One was Pulse
Pressure (PP), which is calculated as
\(\text{PP}=\text{systolic}-\text{diastolic}\). PP is the force of the
hearts contraction on the arteries. The other is called Mean Arterial
Pressure (MAP), which is calculated as
\(\text{MAP}=\frac{\text{systolic}\cdot\text{diastolic}^2}{3}\). MAP is
the average blood pressure throughout a cardiac cycle. We decided on PP
because it was simpler to calculate and understand.

\end{document}

### 2025-05-04 paper/paper.qmd
Update paper.qmd
I added more fluff to our motivation

@@
 
-Sleep is important and people are always wondering how they can improve their sleep.
-Everyone knows how annoying it is to be tossing and turning all night long.
-Lifestyle and general health seem to be important indicators.
-In this report we explore a sleep data set and see what lifestyle and health variables have a significant relationship with sleep duration.
+Sleep is an important contributor to our well-being and is deeply connected to our physical health, and cognitive function.
+However we often find that many individuals struggle with getting long hours sleep, often finding themselves tossing and turning throughout the night.
+Consequently, there is now a growing interest in understanding the lifestyle and bio metric factors that influence sleep patterns and duration.
+General knowledge and intuition suggest factors such as physical activity and occupational stress may influence sleep duration.
+Therefore, our goal is to use linear regression techniques to quantify exactly how much these factors reduce or extend sleep time.
+
+Before, analyzing the data set, we expect a number of our indicators to contribute significantly to sleep duration.
+In general, we would assume that occupation will play an important role in analyzing our data as a more stressful work environment may hinder sleep patterns and disturb the body's natural sleep cycle.
+Additionally, we may expect that higher levels of physical activity will work in tandem with deeper and longer duration of sleep.
+
+Also, we seek to filter through and find the most important variables in our analysis.
+We expect that variables like Daily Steps and Physical Activity Level will be a contribute to sleep duration in a similar manner.
+In our data set, we have numerous combinations of variables that are alike in this way and therefore we want to reduce our factors to give the most precise and simple model possible.
 

### 2025-05-04 Kyubin's Edit/KYUBIN_EDIT.qmd
Create KYUBIN_EDIT.qmd
This is my edit: I did up to LASSO. I going to residual analysis next to see if model has unstable variance which means Ridge and elastic net may be neccesary

---
title: "outlier_exploration"
format: pdf
---

```{r}
library(readr)
library(tidyverse)
library(janitor)

sleep <- read_csv("data/Sleep_health_and_lifestyle_dataset.csv") %>% 
  clean_names()

glimpse(sleep)

sleep <- sleep %>%
  separate(blood_pressure, into = c("systolic", "diastolic"), sep = "/", convert = TRUE)

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$gender <- as.factor(sleep$gender)
sleep$occupation <- as.factor(sleep$occupation)
sleep$bmi_category <- as.factor(sleep$bmi_category)
sleep$sleep_disorder <- as.factor(sleep$sleep_disorder)

sleep <- sleep %>%
  mutate(
    systolic = as.numeric(systolic),
    diastolic = as.numeric(diastolic),
    MAP = (2 * diastolic + systolic) / 3
  )

sleep <- subset(sleep, select = -c(person_id, systolic, diastolic))

glimpse(sleep)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(car)
```

```{r}
full_model <- lm(sleep_duration ~., sleep)
summary(full_model)
```

### Outliers

```{r}
plot(full_model)
```

```{r}
# Calculate Cook's Distance
cooks_d <- cooks.distance(full_model)
summary(cooks_d)

# Threshold
n <- nrow(sleep)
cooks_threshold <- 4 / n

# Basic Cook's Distance plot
plot(cooks_d, type = "h", 
     main = "Cook's Distance", 
     ylab = "Cook's D", 
     xlab = "Observation", 
     col = "blue")

# Add horizontal threshold line
abline(h = cooks_threshold, col = "red", lty = 2)

# Label points above threshold
high_cooks <- which(cooks_d > cooks_threshold)

# Add text labels
text(x = high_cooks, 
     y = cooks_d[high_cooks], 
     labels = high_cooks, 
     pos = 3, 
     cex = 0.8, 
     col = "red")

# Found that obs 6, 7, 107 were pretty large

```

```{r}
# Look into obs 6, 7, 107
sleep[6,]
sleep[7,]
sleep[107,]

# They all seem pretty valid
```

### Multicollinearity

```{r}
library(car)

vif(full_model)

#GVIF high for occupation, stress, bmi category
#GVIF^ high for quality of sleep

#stress
sleep$stress_level_num <- as.numeric(as.factor(sleep$stress_level))

aux_model_stress_level <- lm(stress_level_num ~ gender + age + quality_of_sleep + 
                    physical_activity_level + occupation + 
                    bmi_category + heart_rate + daily_steps + 
                    sleep_disorder + MAP, data = sleep)
summary(aux_model_stress_level)
# R^2 .9542 suggest we remove it

#occupation
sleep$occupation_num <- as.numeric(as.factor(sleep$occupation))

aux_model_occupation <- lm(occupation_num ~ gender + age + quality_of_sleep + 
                           physical_activity_level + bmi_category + heart_rate + 
                           daily_steps + sleep_disorder + MAP + stress_level, 
                           data = sleep)

summary(aux_model_occupation)
# R^2 around .7232 keep occupation

# BMI
sleep$bmi_category_num <- as.numeric(as.factor(sleep$bmi_category))
aux_model_bmi <- lm(bmi_category_num ~ gender + age + quality_of_sleep + 
                     physical_activity_level + occupation + heart_rate + 
                     daily_steps + sleep_disorder + MAP + stress_level,
                     data = sleep)

summary(aux_model_bmi)
# R^2 around 0.9375

#quality
sleep$quality_of_sleep_num <- as.numeric(as.factor(sleep$quality_of_sleep))
aux_model_quality <- lm(quality_of_sleep_num ~ gender + age + physical_activity_level +
                         occupation + bmi_category + heart_rate + daily_steps +
                         sleep_disorder + MAP + stress_level,
                         data = sleep)

summary(aux_model_quality)
# R^2 around 0.9653

```

## Table of R^2 Auxiliary Model
```{r}
library(dplyr)

# Define the full list of predictors (excluding the outcome)
predictors <- c("stress_level", "gender", "age", "quality_of_sleep",
                "physical_activity_level", "occupation", "bmi_category",
                "heart_rate", "daily_steps", "sleep_disorder", "MAP")

# Create a function to fit auxiliary model and return R² and Adjusted R²
get_aux_model_r2 <- function(var, data) {
  # Convert response to numeric (safely)
  data[[paste0(var, "_num")]] <- as.numeric(as.factor(data[[var]]))

  # Define predictors excluding the variable of interest
  others <- setdiff(predictors, var)
  formula <- as.formula(paste(paste0(var, "_num"), "~", paste(others, collapse = " + ")))

  # Fit auxiliary model
  model <- lm(formula, data = data)

  # Extract R² and Adjusted R²
  r2 <- summary(model)$r.squared
  adj_r2 <- summary(model)$adj.r.squared

  return(data.frame(Variable = var, R_squared = r2, Adjusted_R_squared = adj_r2))
}

# Apply the function to all predictors
aux_results <- do.call(rbind, lapply(predictors, get_aux_model_r2, data = sleep))

# Print results
print(aux_results)
```

```{r}

# get rid of 3 highest aux r^2 adj and high vif scores: BMI and Quality
model_vif <- lm(sleep_duration ~gender + age + stress_level +  
                    physical_activity_level + occupation + 
                    heart_rate + daily_steps + 
                    sleep_disorder + MAP, data = sleep)

# View the summary
summary(model_vif)

vif(model_vif)
```

### Variable Selection w VIF
```{r}

# 2. Backward Elimination
model_backward <- step(model_vif, direction = "backward")
summary(model_backward)

# 3. Forward Selection
# Define the full scope and a null model
full_scope <- formula(model_vif)
null_model <- lm(sleep_duration ~ 1, data = sleep)

model_forward <- step(null_model, scope = full_scope, direction = "forward")
summary(model_forward)

# 4. Stepwise Selection (both directions)
model_stepwise <- step(null_model, scope = full_scope, direction = "both")
summary(model_stepwise)

```

# Lasso after VIF and Variable Selection
```{r}

library(glmnet)
library(tidyverse)
library(tidymodels)
library(tidyr)
library(car)
library(caret)

sleep <- subset(sleep, select = -c(bmi_category, quality_of_sleep))

## Split Training/Test Data
sleep_sp <- sleep %>% initial_split(prop = 0.8, strata=sleep_duration)
rtrain <- sleep_sp %>% training()
rtest <- sleep_sp %>% testing()

model0 <- lm(sleep_duration ~ ., data=rtrain)
summary(model0)

## Regularization
xrtrain <- rtrain %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtrain <- rtrain$sleep_duration

xrtest  <- rtest %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtest  <- rtest$sleep_duration

modelL <- glmnet(xrtrain, yrtrain)
plot(modelL,label = TRUE)
## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
print(modelL)
## Coefficient for a certain lambda
coef.glmnet(modelL, s= 0.1)

## Cross-Validation
## nfolds = # of groups ex) 20
## Regression with 19 groups and test with 1 group
cvmodelL <- cv.glmnet(xrtrain, yrtrain, type.measure = "mse", nfolds = 20)
coef.glmnet(modelL, s= cvmodelL$lambda.min)
plot(cvmodelL)

coef_min_lambda <- coef.glmnet(modelL, s = cvmodelL$lambda.min)

# Convert to a tidy data frame and filter out zero coefficients
nonzero_coefs <- as.matrix(coef_min_lambda)
nonzero_coefs_df <- data.frame(
  Variable = rownames(nonzero_coefs),
  Coefficient = nonzero_coefs[, 1]
)

# View coefficients that were shrunk to zero
zeroed_out <- nonzero_coefs_df %>% filter(Coefficient == 0)
print(zeroed_out)

# gender, occupation and sleep disorder zeroed out and age is insignificant

model_LASSO <- lm(sleep_duration ~ stress_level +  
                    physical_activity_level + 
                    heart_rate + daily_steps + 
                    MAP, data = sleep)
summary(model_LASSO)

```
## Ridge if neccessary

```{r}
modelR <- glmnet(xrtrain, yrtrain, alpha=0)
plot(modelR,label = TRUE)
## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
print(modelR)
## Coefficient for a certain lambda
coef.glmnet(modelR, s= 0.1) ## s= lambda value (any)

## CrossValidation
cvmodelR <- cv.glmnet(xrtrain, yrtrain, alpha=0, type.measure = "mse", nfolds = 20)
plot(cvmodelR)
log(cvmodelR$lambda.min)
```

### 2025-05-05 paper/paper.qmd
adding more stuff to the paper

--- 2025-05-03 paper/paper.qmd
@@
 library(janitor)
+library(patchwork)
+library(broom)
 ```
@@
 Gender, Occupation, Quality of Sleep, Stress Level, BMI Category and Sleep Disorder are categorical variables.
-Physical Activity, MAP, and Daily steps are continuous variables.
+Physical Activity, PP, agem and Daily steps are continuous variables.
 We have a total of 374 observations.
@@
 We decided on PP because it was simpler to calculate and understand.
+
+For transparency it should be noted that the data is synthetic. Even though it is synthetic we still treat the analysis and report as if it was real data.
+
+\vspace{1.5em}
+
+\noindent\textbf{\Large Exploratory Data Analysis}
+
+```{r}
+p1 <- sleep %>% 
+  ggplot() +
+  geom_jitter(aes(x = age, y = sleep_duration), size = .5) +
+  theme_few() +
+  labs(x = "Age", y = NULL)
+
+p2 <- sleep %>% 
+  ggplot() +
+  geom_jitter(aes(x = PP, y = sleep_duration), size = .5) +
+  theme_few() +
+  labs(x = "Pulse Pressure", y = NULL)
+
+p3 <- sleep %>% 
+  ggplot() +
+  geom_jitter(aes(x = heart_rate, y = sleep_duration), size = .5) +
+  theme_few() +
+  labs(x = "Heart Rate", y = NULL)
+
+p4 <- sleep %>% 
+  ggplot() +
+  geom_jitter(aes(x = daily_steps, y = sleep_duration), size = .5) +
+  theme_few() +
+  labs(x = "Daily Steps", y = NULL)
+
+(p1 + p2) / (p3 + p4) +
+  plot_annotation(
+    title = "Sleep Duration vs Continuous Variables",
+    theme = theme(plot.title = element_text(size = 16, hjust = 0.5))
+  ) &
+  theme(axis.title.y = element_text(angle = 90)) &
+  labs(y = "Sleep Duration")
+```
+
+```{r}
+corr_vars <- sleep %>%
+  dplyr::select(sleep_duration, age, PP, heart_rate, daily_steps)
+
+corr_matrix <- round(cor(corr_vars, use = "complete.obs"), 2)
+
+knitr::kable(corr_matrix, caption = "Correlation Matrix of Continuous Varaibles")
+```
+
+\vspace{1.5em}
+
+\noindent\textbf{\Large Full Model}
+
+Running the full model we can see that our omnibus hypothesis test has an F-stat of 426.6 which is greater than the 99th percentile for our F-distribution, $F_{31,342,.99}=$ `r round(qf(0.99, 31, 342),2)`.
+So we reject the null that all coefficients are 0.
+We also see a high Adjusted $R^2$ and very low Residual Sum of Squares compared to the Total Sum of Squares.
+
+```{r}
+# Fit the model
+full_model <- lm(sleep_duration ~ ., data = sleep)
+
+# Get summary
+model_summary <- summary(full_model)
+
+model_stats <- tibble::tibble(`F-statistic` = round(model_summary$fstatistic[1], 2),
+                              `Adjusted R²` = round(model_summary$adj.r.squared, 3),
+                              `Residual Sum of Squares` = round(sum(residuals(full_model)^2), 2),
+                              `Total Sum of Squares` = round(sum((sleep$sleep_duration - mean(sleep$sleep_duration))^2), 2))
+
+knitr::kable(model_stats, caption = "Full Linear Model Outcome")
+```
+
+As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but their is no excessive deviation from normality.
+
+```{r}
+# Extract model diagnostics
+model_data <- augment(full_model)
+
+# Residuals vs Fitted plot
+p5 <- ggplot(model_data, aes(.fitted, .resid)) +
+  geom_point(size = 1) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  theme_few() +
+  labs(title = "Residuals vs Fitted",
+       x = "Fitted values", y = "Residuals")
+
+# Q-Q plot
+p6 <- ggplot(model_data, aes(sample = .resid)) +
+  stat_qq(size = 1) +
+  stat_qq_line() +
+  theme_few() +
+  labs(title = "Normal Q-Q Plot", x = "Theoretical Quantiles", y = "Sample Quantiles")
+
+# Combine with patchwork
+p5 / p6
+```

### 2025-05-05 paper/paper.qmd
outlier add

@@
 
-\vspace{1.5em}
+\newpage
 
@@
 ```{r}
-# Fit the model
 full_model <- lm(sleep_duration ~ ., data = sleep)
 
-# Get summary
 model_summary <- summary(full_model)
@@
 
-As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but their is no excessive deviation from normality.
-
 ```{r}
-# Extract model diagnostics
 model_data <- augment(full_model)
 
-# Residuals vs Fitted plot
 p5 <- ggplot(model_data, aes(.fitted, .resid)) +
@@
 
-# Q-Q plot
 p6 <- ggplot(model_data, aes(sample = .resid)) +
@@
 
-# Combine with patchwork
 p5 / p6
 ```
+
+As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
+
+Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw

### 2025-05-05 paper/paper.qmd
working outlier and multicoll

@@
 full_model <- lm(sleep_duration ~ ., data = sleep)
+full_model_summary <- summary(full_model)
 
-model_summary <- summary(full_model)
-
-model_stats <- tibble::tibble(`F-statistic` = round(model_summary$fstatistic[1], 2),
-                              `Adjusted R²` = round(model_summary$adj.r.squared, 3),
+model_stats <- tibble::tibble(`F-statistic` = round(full_model_summary$fstatistic[1], 2),
+                              `Adjusted R²` = round(full_model_summary$adj.r.squared, 3),
                               `Residual Sum of Squares` = round(sum(residuals(full_model)^2), 2),
@@
 
+```{r}
+kable(vif(full_model))
+```
+
 Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw

### 2025-05-06 paper/paper.qmd
paper work

@@
 sleep$stress_level <- as.factor(sleep$stress_level)
-sleep$quality_of_sleep <- as.factor(sleep$quality_of_sleep)
+sleep$gender <- as.factor(sleep$gender)
+sleep$occupation <- as.factor(sleep$occupation)
+sleep$bmi_category <- as.factor(sleep$bmi_category)
+sleep$sleep_disorder <- as.factor(sleep$sleep_disorder)
 
@@
 ```
+We also saw all coefficients have standard error less than one.
 
@@
 As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
+We will explore multicollinearity and outliers next.
 
@@
 ```
+`age` and `quality_of_sleep` have $\text{GVIF}^{2\cdot\text{DF}} > 5$. We may want to explore Lasso to deal with this.
 
 Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw
+
+```{r}
+#Values of Interest
+leverage <- hatvalues(full_model)
+studentized <- rstudent(full_model)
+dffits_vals <- dffits(full_model)
+cooks_d <- cooks.distance(full_model)
+
+diagnostics <- data.frame(Leverage = leverage,
+                          Studentized = studentized,
+                          DFFITS = dffits_vals,
+                          Cooks_Distance = cooks_d)
+
+#Thresh hold
+n <- nrow(sleep)
+p <- length(coef(full_model)) - 1
+leverage_threshold <- 2 * (p + 1) / n
+studentized_threshold <- 2
+cooks_threshold <- 1
+
+# Find points exceeding thresholds
+high_leverage <- which(leverage > leverage_threshold)
+high_studentized <- which(abs(studentized) > studentized_threshold)
+high_cooks <- which(cooks_d > cooks_threshold)
+
+# Unique indices that exceed any threshold
+suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))
+
+sleep[suspicious_indices,]
+```
+
+```{r}
+leverage <- hatvalues(full_model)
+studentized <- rstudent(full_model)
+dffits_vals <- dffits(full_model)
+cooks_d <- cooks.distance(full_model)
+
+diagnostics <- data.frame(Leverage = leverage,
+                          Studentized = studentized,
+                          DFFITS = dffits_vals,
+                          Cooks_Distance = cooks_d)
+
+n <- nrow(sleep)
+p <- length(coef(full_model)) - 1
+leverage_threshold <- 2 * (p + 1) / n
+studentized_threshold <- 2
+cooks_threshold <- 1
+
+# Find points exceeding thresholds
+high_leverage <- which(leverage > leverage_threshold)
+high_studentized <- which(abs(studentized) > studentized_threshold)
+high_cooks <- which(cooks_d > cooks_threshold)
+
+# Unique indices that exceed any threshold
+suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))
+
+# Return threshold values, diagnostic preview, and rows flagged as potential outliers
+list(
+  leverage_threshold = leverage_threshold,
+  studentized_threshold = studentized_threshold,
+  cooks_threshold = cooks_threshold,
+  diagnostics_preview = head(diagnostics, 10),
+  flagged_rows = diagnostics[suspicious_indices, ])
+
+# Remove the suspicious rows
+sleep_clean <- sleep[-suspicious_indices, ]
+
+# Refit the model excluding suspicious rows 
+full_model_clean <- lm(sleep_duration ~ ., data = sleep_clean)
+
+# Plot diagnostics again on the cleaned model
+par(mfrow = c(2, 2))
+plot(full_model_clean)
+
+# Comparison of model summaries
+summary(full_model)        # Old model (with potential outliers)  R^2  = 0.9592
+summary(full_model_clean)  # New model (without potential outliers)  R^2  = 0.9929
+
+```

### 2025-05-06 paper/paper.qmd
commit

@@
 ```
+Visual relationship can be seen for `Sleep Duration` with `Age` and `sleep_duration`.
+Not so much for `PP` and `daily_steps`.
+The correlation between our numeric covariates is most notable between `sleep_duration` and `heart_rate`.
+`age` and `hearts_pressure`
 
@@
 
-Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw
+\newpage
+
+Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1.
+We saw a total of 57 observations that exceeded at least one of these thresholds.
+Since we can't say that any of these observation are errors we will only remove row 264 because it has a leverage of 1 and `NaN` for the rest.
+This is clearly a problem observation.
 
@@
 
-sleep[suspicious_indices,]
+aggregation <- list(leverage_threshold = leverage_threshold,
+                    studentized_threshold = studentized_threshold,
+                    cooks_threshold = cooks_threshold,
+                    flagged_rows = diagnostics[suspicious_indices, ])
 ```
@@
 ```{r}
-leverage <- hatvalues(full_model)
-studentized <- rstudent(full_model)
-dffits_vals <- dffits(full_model)
-cooks_d <- cooks.distance(full_model)
+aggregation$flagged_rows %>% 
+  arrange(desc(Leverage)) %>% 
+  head(n = 1) %>% 
+  kable(digits = 3, caption = "Row with Highest Leverage")
+```
 
-diagnostics <- data.frame(Leverage = leverage,
-                          Studentized = studentized,
-                          DFFITS = dffits_vals,
-                          Cooks_Distance = cooks_d)
+\noindent\textbf{\Large Updated Model}
 
-n <- nrow(sleep)
-p <- length(coef(full_model)) - 1
-leverage_threshold <- 2 * (p + 1) / n
-studentized_threshold <- 2
-cooks_threshold <- 1
-
-# Find points exceeding thresholds
-high_leverage <- which(leverage > leverage_threshold)
-high_studentized <- which(abs(studentized) > studentized_threshold)
-high_cooks <- which(cooks_d > cooks_threshold)
-
-# Unique indices that exceed any threshold
-suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))
-
-# Return threshold values, diagnostic preview, and rows flagged as potential outliers
-list(
-  leverage_threshold = leverage_threshold,
-  studentized_threshold = studentized_threshold,
-  cooks_threshold = cooks_threshold,
-  diagnostics_preview = head(diagnostics, 10),
-  flagged_rows = diagnostics[suspicious_indices, ])
-
-# Remove the suspicious rows
-sleep_clean <- sleep[-suspicious_indices, ]
-
-# Refit the model excluding suspicious rows 
-full_model_clean <- lm(sleep_duration ~ ., data = sleep_clean)
-
-# Plot diagnostics again on the cleaned model
-par(mfrow = c(2, 2))
-plot(full_model_clean)
-
-# Comparison of model summaries
-summary(full_model)        # Old model (with potential outliers)  R^2  = 0.9592
-summary(full_model_clean)  # New model (without potential outliers)  R^2  = 0.9929
-
+```{r}
+sleep_test <- sleep[-264,]
 ```

### 2025-05-07 data/data_wrangling.R
hi

@@ -19,7 +19,7 @@ sleep <- sleep %>%
     MAP = (2 * diastolic + systolic) / 3
   )
 
-sleep <- sleep %>% 
-  select(!c(person_id, systolic, diastolic))
+sleep <- subset(sleep, select = -c(person_id, systolic, diastolic))
+
 
 glimpse(sleep)

### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
Changed the data wrangling in KYUBIN_EDIT to the new one, incorporating "pp" variable

@@
 library(janitor)
+library(dplyr)
 
-sleep <- read_csv("data/Sleep_health_and_lifestyle_dataset.csv") %>% 
+sleep <- read_csv("../data/Sleep_health_and_lifestyle_dataset.csv") %>%  
   clean_names()
-
-glimpse(sleep)
 
@@
 sleep <- sleep %>%
-  mutate(
-    systolic = as.numeric(systolic),
+  mutate(systolic = as.numeric(systolic),
          diastolic = as.numeric(diastolic),
-    MAP = (2 * diastolic + systolic) / 3
-  )
+         PP = systolic-diastolic)
 
-sleep <- subset(sleep, select = -c(person_id, systolic, diastolic))
-
-glimpse(sleep)
+sleep <- sleep %>%
+  dplyr::select(!c(person_id, systolic, diastolic))
 

### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
Lines 96-194 were edited: Added in the Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier Created the new dataset that removed obs 264 called sleep_test Created the new full model that removed obs 264 called full_model_test Changed "sleep" to "sleep_test" for for the VIF testing

@@
 
+### Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier
+
+```{r}
+leverage <- hatvalues(full_model)
+studentized <- rstudent(full_model)
+dffits_vals <- dffits(full_model)
+cooks_d <- cooks.distance(full_model)
+
+diagnostics <- data.frame(Leverage = leverage,
+                          Studentized = studentized,
+                          DFFITS = dffits_vals,
+                          Cooks_Distance = cooks_d)
+
+n <- nrow(sleep)
+p <- length(coef(full_model)) - 1
+leverage_threshold <- 2 * (p + 1) / n
+studentized_threshold <- 2
+cooks_threshold <- 1
+
+# Find points exceeding thresholds
+high_leverage <- which(leverage > leverage_threshold)
+high_studentized <- which(abs(studentized) > studentized_threshold)
+high_cooks <- which(cooks_d > cooks_threshold)
+
+# Unique indices that exceed any threshold
+suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))
+
+# Return threshold values, diagnostic preview, and rows flagged as potential outliers
+list(
+  leverage_threshold = leverage_threshold,
+  studentized_threshold = studentized_threshold,
+  cooks_threshold = cooks_threshold,
+  diagnostics_preview = head(diagnostics, 10),
+  flagged_rows = diagnostics[suspicious_indices, ])
+
+# observation 264 has a leverage of 1
+
+# New data set used for model
+sleep_test <- sleep[-264,]
+
+full_model_test<-lm(sleep_duration ~., sleep_test)
+summary(full_model_test)
+```
+
 ### Multicollinearity
@@
 
-vif(full_model)
+vif(full_model_test)
 
-#GVIF high for occupation, stress, bmi category
+#GVIF high for occupation, stress level, bmi category
 #GVIF^ high for quality of sleep
@@
 #stress
-sleep$stress_level_num <- as.numeric(as.factor(sleep$stress_level))
+sleep_test$stress_level_num <- as.numeric(as.factor(sleep_test$stress_level))
 
@@
                     bmi_category + heart_rate + daily_steps + 
-                    sleep_disorder + MAP, data = sleep)
+                    sleep_disorder + PP, data = sleep_test)
 summary(aux_model_stress_level)
-# R^2 .9542 suggest we remove it
+# R^2 .9563 suggest we remove it
 
 #occupation
-sleep$occupation_num <- as.numeric(as.factor(sleep$occupation))
+sleep_test$occupation_num <- as.numeric(as.factor(sleep_test$occupation))
 
@@
                            physical_activity_level + bmi_category + heart_rate + 
-                           daily_steps + sleep_disorder + MAP + stress_level, 
-                           data = sleep)
+                           daily_steps + sleep_disorder + PP + stress_level, 
+                           data = sleep_test)
 
 summary(aux_model_occupation)
-# R^2 around .7232 keep occupation
+# R^2 around .7453, keep occupation
 
 # BMI
-sleep$bmi_category_num <- as.numeric(as.factor(sleep$bmi_category))
+sleep_test$bmi_category_num <- as.numeric(as.factor(sleep_test$bmi_category))
+
 aux_model_bmi <- lm(bmi_category_num ~ gender + age + quality_of_sleep + 
                      physical_activity_level + occupation + heart_rate + 
-                     daily_steps + sleep_disorder + MAP + stress_level,
-                     data = sleep)
+                     daily_steps + sleep_disorder + PP + stress_level,
+                     data = sleep_test)
 
 summary(aux_model_bmi)
-# R^2 around 0.9375
+# R^2 around 0.9326
 
 #quality
-sleep$quality_of_sleep_num <- as.numeric(as.factor(sleep$quality_of_sleep))
+sleep_test$quality_of_sleep_num <- as.numeric(as.factor(sleep_test$quality_of_sleep))
+
 aux_model_quality <- lm(quality_of_sleep_num ~ gender + age + physical_activity_level +
                          occupation + bmi_category + heart_rate + daily_steps +
-                         sleep_disorder + MAP + stress_level,
-                         data = sleep)
+                         sleep_disorder + PP + stress_level,
+                         data = sleep_test)
 
 summary(aux_model_quality)
-# R^2 around 0.9653
+# R^2 around 0.9673
 

### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
lines 200 - 266 Continued replacing "sleep" with "sleep_test" Replaced "MAP" with "PP" variable in the predictors vector (line 203) After the VIF function is ran, the two variables with the highest R^2 Adj were Age and Quality of Sleep, so they were removed in the model_vif

@@
 
-# observation 264 has a leverage of 1
+# observation 264 has a leverage of 1 !
 
@@
                 "physical_activity_level", "occupation", "bmi_category",
-                "heart_rate", "daily_steps", "sleep_disorder", "MAP")
+                "heart_rate", "daily_steps", "sleep_disorder", "PP")
 
@@
 # Apply the function to all predictors
-aux_results <- do.call(rbind, lapply(predictors, get_aux_model_r2, data = sleep))
+aux_results <- do.call(rbind, lapply(predictors, get_aux_model_r2, data = sleep_test))
 
@@
 
-# get rid of 3 highest aux r^2 adj and high vif scores: BMI and Quality
-model_vif <- lm(sleep_duration ~gender + age + stress_level +  
+# get rid of 3 highest aux r^2 adj and high VIF scores: Age and Quality of sleep
+model_vif <- lm(sleep_duration ~gender + bmi_category + stress_level +  
                     physical_activity_level + occupation + 
                     heart_rate + daily_steps + 
-                    sleep_disorder + MAP, data = sleep)
+                    sleep_disorder + PP, data = sleep_test)
 

### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
line 275 Added a message that explains where the editing was left off at

@@
 
-# get rid of 3 highest aux r^2 adj and high VIF scores: Age and Quality of sleep
-model_vif <- lm(sleep_duration ~gender + bmi_category + stress_level +  
+# get rid of 2 highest aux r^2 adj and high VIF scores: Age and Quality of Sleep
+model_vif <- lm(sleep_duration ~gender + bmi_category  +
                     physical_activity_level + occupation + 
                     heart_rate + daily_steps + 
-                    sleep_disorder + PP, data = sleep_test)
+                    sleep_disorder + PP + stress_level, data = sleep_test)
 
@@
 full_scope <- formula(model_vif)
-null_model <- lm(sleep_duration ~ 1, data = sleep)
+null_model <- lm(sleep_duration ~ 1, data = sleep_test)
 
@@
 ```{r}
+
+#### ALL THE CODE BELOW THIS HAS NOT BEEN EDITED! STILL NEED TO CHANGE "sleep" TO "sleep_test" AND CHANGE "bmi_category" TO "age" ####
 

### 2025-05-07 paper/paper.qmd
word change

@@
 
-As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
+As far as assumptions go homoskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
 We will explore multicollinearity and outliers next.

### 2025-05-07 Kyubin's Edit/paper_KYUBIN.qmd
Create paper_KYUBIN.qmd

--- 2025-05-07 paper/paper.qmd
@@
 
-```{r}
-kable(vif(full_model))
-```
-`age` and `quality_of_sleep` have $\text{GVIF}^{2\cdot\text{DF}} > 5$. We may want to explore Lasso to deal with this.
+\noindent\textbf{\Large Regression Diagnostics: Outliers and Multicollinearity}
 
-\newpage
-
-Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1.
+Looking now towards outliers, we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1.
 We saw a total of 57 observations that exceeded at least one of these thresholds.
@@
 
-\noindent\textbf{\Large Updated Model}
+Next, we want to take care of multicollinearity after removing observation 264 from the data set.
+We checked the VIF values for each input covariate and determined which features can be removed.
+We also set 5 as a threshold for a large VIF value and removed any features that exceed this value.
+Furthermore, we used the scaled GVIF value, $\text{GVIF}^{1 / (2 \cdot \text{Df})}$, as it allows categorical variables to be numerically compared to continuous variables.
+From our VIF table, we noticed that `age` and `quality_of_sleep` have $\text{GVIF}^{1 / (2 \cdot \text{Df})} > 5$.
 
@@
 sleep_test <- sleep[-264,]
+
+full_model_test<-lm(sleep_duration ~., sleep_test)
+summary(full_model_test)
+
+library(car)
+
+kable(vif(full_model_test))
 ```
+
+We removed variables `age` and `quality_of_sleep` and ran a MLR model.
+Our results showed that the results from reduced VIF model was in very similar to the full model.
+Adjusted R^2 was essentially unchanged.
+Therefore, we concluded that `age` and `quality_of_sleep` did not contribute much to the regression analysis because they were linear combinations of other independent variables.
+
+```{r}
+sleep_test_vif <- sleep_test %>%
+  dplyr::select(!c(age, quality_of_sleep))
+
+full_model_vif <- lm(sleep_duration ~., sleep_test_vif)
+
+model_stats_vif <- tibble::tibble(
+  `F-stat.` = round(summary(full_model_vif)$fstatistic[1], 2),
+  `Adjusted R²` = round(summary(full_model_vif)$adj.r.squared, 3),
+  `Residual Sum of Squares` = round(sum(residuals(full_model_vif)^2), 2),
+  `Total Sum of Squares` = round(sum((sleep_test_vif$sleep_duration - mean(sleep_test_vif$sleep_duration))^2), 2))
+
+knitr::kable(model_stats_vif, caption = "Outlier + VIF Model")
+
+```

### 2025-05-08 Kyubin's Edit/paper_KYUBIN.qmd

@@
 
+\vspace{0.5em}
+
 Running the full model we can see that our omnibus hypothesis test has an F-stat of 426.6 which is greater than the 99th percentile for our F-distribution, $F_{31,342,.99}=$ `r round(qf(0.99, 31, 342),2)`.
@@
 
+\vspace{1.5em}
+
 \noindent\textbf{\Large Regression Diagnostics: Outliers and Multicollinearity}
+
+\vspace{0.5em}
 
@@
 
-Next, we want to take care of multicollinearity after removing observation 264 from the data set.
+\newpage
+
+Next, we wanted to take care of multicollinearity after removing observation 264 from the data set.
 We checked the VIF values for each input covariate and determined which features can be removed.
@@
 
+\vspace{0.5em}
+
 ```{r}
@@
 full_model_test<-lm(sleep_duration ~., sleep_test)
-summary(full_model_test)
 
@@
 
+\vspace{0.5em}
+
 We removed variables `age` and `quality_of_sleep` and ran a MLR model.
-Our results showed that the results from reduced VIF model was in very similar to the full model.
+Our results showed that the reduced VIF model was very similar to the full model.
 Adjusted R^2 was essentially unchanged.
@@
 
-```{r}
+\vspace{0.5em}
+
+```{r include=FALSE}
+
 sleep_test_vif <- sleep_test %>%
@@
 
+```
+
+```{r}
 model_stats_vif <- tibble::tibble(
@@
 knitr::kable(model_stats_vif, caption = "Outlier + VIF Model")
+```
+
+\vspace{1.5em}
+
+\noindent\textbf{\Large Variable Selection: Backward, Forward, and Stepwise}
+
+\vspace{0.5em}
+
+In order to avoid overfitting issues, we utilized a greedy algorithmn process to select variables that provided the lowest AIC value.
+We ran 3 model selection process as discussed in class and recorded the variables selected for each method.
+We found that all 3 methods took all of our variables in our data set.
+We knew that AIC penalizes the fitted model for each additional variables taken.
+Despite this, we saw that all variables were significant.
+Note that the variable selection process was done after our high VIF variables were removed.
+
+```{r include=FALSE}
+# Load required packages
+library(knitr)
+
+# 1. Model Selection using full_model_vif
+model_backward <- step(full_model_vif, direction = "backward")
+
+# Define full model formula and null model for forward/stepwise
+full_scope <- formula(full_model_vif)
+null_model <- lm(sleep_duration ~ 1, data = sleep)
+
+model_forward <- step(null_model, scope = full_scope, direction = "forward")
+model_stepwise <- step(null_model, scope = full_scope, direction = "both")
+
+# 2. Extract variable names selected by each model
+vars_backward <- attr(terms(model_backward), "term.labels")
+vars_forward <- attr(terms(model_forward), "term.labels")
+vars_stepwise <- attr(terms(model_stepwise), "term.labels")
+
+# 3. Combine all variables that appeared in any model
+all_vars <- sort(unique(c(vars_backward, vars_forward, vars_stepwise)))
+
+# 4. Create a comparison table
+selection_table <- data.frame(
+  Variable = all_vars,
+  Backward = ifelse(all_vars %in% vars_backward, "Yes", "No"),
+  Forward  = ifelse(all_vars %in% vars_forward, "Yes", "No"),
+  Stepwise = ifelse(all_vars %in% vars_stepwise, "Yes", "No")
+)
 
 ```
+
+```{r}
+
+kable(selection_table, caption = "Variable Selection")
+
+```
+
+\vspace{1.5em}
+
+\noindent\textbf{\Large Suggested Regression: LASSO}
+
+\vspace{0.5em}
+
+For our suggested regression, we used LASSO regression.
+We hoped to extract the most import input features to achieve a simple yet powerful model.
+To find the optimal $\lambda$ or the regularization parameter, we used a cross-validation process by splitting our data set: train = 80%.
+We saw that `gender`, `occupation`, `bmi_category`, and `sleep_disorder` converged to zero.
+
+```{r include=FALSE}
+
+library(glmnet)
+library(tidyverse)
+library(tidymodels)
+library(tidyr)
+library(car)
+library(caret)
+
+## Split Training/Test Data
+sleep_sp <- sleep_test_vif %>% initial_split(prop = 0.8, strata=sleep_duration)
+rtrain <- sleep_sp %>% training()
+rtest <- sleep_sp %>% testing()
+
+model0 <- lm(sleep_duration ~ ., data=rtrain)
+summary(model0)
+
+## Regularization
+xrtrain <- rtrain %>% dplyr::select(-sleep_duration) %>% as.matrix()
+yrtrain <- rtrain$sleep_duration
+
+xrtest  <- rtest %>% dplyr::select(-sleep_duration) %>% as.matrix()
+yrtest  <- rtest$sleep_duration
+
+modelL <- glmnet(xrtrain, yrtrain)
+plot(modelL,label = TRUE)
+## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
+print(modelL)
+## Coefficient for a certain lambda
+coef.glmnet(modelL, s= 0.1)
+
+## Cross-Validation
+## nfolds = # of groups ex) 20
+## Regression with 19 groups and test with 1 group
+cvmodelL <- cv.glmnet(xrtrain, yrtrain, type.measure = "mse", nfolds = 20)
+coef.glmnet(modelL, s= cvmodelL$lambda.min)
+plot(cvmodelL)
+
+coef_min_lambda <- coef.glmnet(modelL, s = cvmodelL$lambda.min)
+
+# Convert to a tidy data frame and filter out zero coefficients
+nonzero_coefs <- as.matrix(coef_min_lambda)
+nonzero_coefs_df <- data.frame(
+  Variable = rownames(nonzero_coefs),
+  Coefficient = nonzero_coefs[, 1]
+)
+
+```
+
+```{r}
+kable(nonzero_coefs_df, digits = 4, caption = "Non-Zero Coefficients from LASSO (Selected by Cross-Validation)")
+```

### 2025-05-08 paper/paper.qmd
spelling

@@
 
-As far as assumptions go homoskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
+As far as assumptions go homoscedasticity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.
 We will explore multicollinearity and outliers next.

### 2025-05-08 paper/paper.qmd
adding kyubins stuff

@@
 
-\newpage
-
 Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1.
@@
 
+We removed variables `age` and `quality_of_sleep` and ran a MLR model.
+Our results showed that the reduced VIF model was very similar to the full model.
+Adjusted R^2 was essentially unchanged.
+Therefore, we concluded that `age` and `quality_of_sleep` did not contribute much to the regression analysis because they were linear combinations of other independent variables.
+
 ```{r}
 sleep_test <- sleep[-264,]
+
+sleep_test <- sleep_test %>%
+  dplyr::select(!c(age, quality_of_sleep))
+
+test_model<-lm(sleep_duration ~., sleep_test)
+test_model_summary <- summary(test_model)
+
+test_model_stats <- tibble::tibble(`F-statistic` = round(test_model_summary$fstatistic[1], 2),
+                              `Adjusted R²` = round(test_model_summary$adj.r.squared, 3),
+                              `Residual Sum of Squares` = round(sum(residuals(test_model)^2), 2),
+                              `Total Sum of Squares` = round(sum((sleep_test$sleep_duration-mean(sleep_test$sleep_duration))^2), 2))
+
+knitr::kable(test_model_stats, caption = "MLR Outcome for")
 ```
+
+In order to avoid overfitting issues, we utilized a greedy algorithmn process to select variables that provided the lowest AIC value.
+We ran 3 model selection process as discussed in class and recorded the variables selected for each method.
+We found that all 3 methods took all of our variables in our data set.
+We knew that AIC penalizes the fitted model for each additional variables taken.
+Despite this, we saw that all variables were significant.
+Note that the variable selection process was done after our high VIF variables were removed.
+
+```{r include=FALSE}
+# 1. Model Selection using full_model_vif
+model_backward <- step(test_model, direction = "backward")
+
+# Define full model formula and null model for forward/stepwise
+full_scope <- formula(test_model)
+null_model <- lm(sleep_duration ~ 1, data = sleep_test)
+
+model_forward <- step(null_model, scope = full_scope, direction = "forward")
+model_stepwise <- step(null_model, scope = full_scope, direction = "both")
+
+# 2. Extract variable names selected by each model
+vars_backward <- attr(terms(model_backward), "term.labels")
+vars_forward <- attr(terms(model_forward), "term.labels")
+vars_stepwise <- attr(terms(model_stepwise), "term.labels")
+
+# 3. Combine all variables that appeared in any model
+all_vars <- sort(unique(c(vars_backward, vars_forward, vars_stepwise)))
+
+# 4. Create a comparison table
+selection_table <- data.frame(
+  Variable = all_vars,
+  Backward = ifelse(all_vars %in% vars_backward, "Yes", "No"),
+  Forward  = ifelse(all_vars %in% vars_forward, "Yes", "No"),
+  Stepwise = ifelse(all_vars %in% vars_stepwise, "Yes", "No")
+)
+```
+
+```{r}
+kable(selection_table, caption = "Variable Selection")
+```

### 2025-05-08 paper/paper.qmd
commit

@@
 library(broom)
+library(glmnet)
+library(caret)
+library(tidymodels)
+library(rsample)
 ```
@@
 # 1. Model Selection using full_model_vif
-model_backward <- step(test_model, direction = "backward")
+model_backward <- stats::step(test_model, direction = "backward")
 
@@
 
-model_forward <- step(null_model, scope = full_scope, direction = "forward")
-model_stepwise <- step(null_model, scope = full_scope, direction = "both")
+model_forward <- stats::step(null_model, scope = full_scope, direction = "forward")
+model_stepwise <- stats::step(null_model, scope = full_scope, direction = "both")
 
@@
 ```
+
+For our suggested regression, we used LASSO regression.
+We hoped to extract the most import input features to achieve a simple yet powerful model.
+To find the optimal $\lambda$ or the regularization parameter, we used a cross-validation process by splitting our data set: train = 80%.
+We saw that `gender`, `occupation`, `bmi_category`, and `sleep_disorder` converged to zero.
+
+```{r include=FALSE}
+## Split Training/Test Data
+sleep_sp <- sleep_test %>% initial_split(prop = 0.8, strata=sleep_duration)
+rtrain <- sleep_sp %>% training()
+rtest <- sleep_sp %>% testing()
+
+model0 <- lm(sleep_duration ~ ., data=rtrain)
+summary(model0)
+
+## Regularization
+xrtrain <- rtrain %>% dplyr::select(-sleep_duration) %>% as.matrix()
+yrtrain <- rtrain$sleep_duration
+
+xrtest  <- rtest %>% dplyr::select(-sleep_duration) %>% as.matrix()
+yrtest  <- rtest$sleep_duration
+
+modelL <- glmnet(xrtrain, yrtrain)
+plot(modelL,label = TRUE)
+## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
+print(modelL)
+## Coefficient for a certain lambda
+coef.glmnet(modelL, s= 0.1)
+
+## Cross-Validation
+## nfolds = # of groups ex) 20
+## Regression with 19 groups and test with 1 group
+cvmodelL <- cv.glmnet(xrtrain, yrtrain, type.measure = "mse", nfolds = 20)
+coef.glmnet(modelL, s= cvmodelL$lambda.min)
+plot(cvmodelL)
+
+coef_min_lambda <- coef.glmnet(modelL, s = cvmodelL$lambda.min)
+
+# Convert to a tidy data frame and filter out zero coefficients
+nonzero_coefs <- as.matrix(coef_min_lambda)
+nonzero_coefs_df <- data.frame(
+  Variable = rownames(nonzero_coefs),
+  Coefficient = nonzero_coefs[, 1]
+)
+
+```
+
+```{r}
+kable(nonzero_coefs_df, digits = 4, caption = "Non-Zero Coefficients from LASSO (Selected by Cross-Validation)",
+      row.names = FALSE)
+```

### 2025-05-08 paper/paper.qmd
set seed 510 and final LASSO regression model

@@
 ```{r include=FALSE}
+set.seed(510)
+
 ## Split Training/Test Data
@@
 ```
+
+```{r}
+sleep_test_final <- sleep_test_vif %>%
+  dplyr::select(!c(gender, occupation, bmi_category, sleep_disorder, daily_steps))
+
+model_final <- lm(sleep_duration ~., sleep_test_final)
+summary(model_final)
+```

### 2025-05-08 paper/paper.qmd
kable

@@
 
-knitr::kable(test_model_stats, caption = "MLR Outcome for")
+knitr::kable(test_model_stats, caption = "MLR Outcome for age and quality of sleep removed")
 ```
@@
 ```{r}
-sleep_test_final <- sleep_test_vif %>%
+sleep_test_final <- sleep_test %>%
   dplyr::select(!c(gender, occupation, bmi_category, sleep_disorder, daily_steps))
 
-model_final <- lm(sleep_duration ~., sleep_test_final)
-summary(model_final)
+final_model <- lm(sleep_duration ~., sleep_test_final)
+final_model_summary <- summary(model_final)
+
+final_model_stats <- tibble::tibble(`F-statistic` = round(final_model_summary$fstatistic[1], 2),
+                              `Adjusted R²` = round(final_model_summary$adj.r.squared, 3),
+                              `Residual Sum of Squares` = round(sum(residuals(final_model)^2), 2),
+                              `Total Sum of Squares` = round(sum((sleep_test_final$sleep_duration-mean(sleep_test_final$sleep_duration))^2), 2))
+
+knitr::kable(final_model_stats, caption = "Final Model Statistics")
 ```

### 2025-05-08 paper/paper.qmd
fix
formatting
cross validation

@@
 
-\noindent\textbf{\Large Motivation}
+\begin{center}
+\Large\textbf{\large Introduction}
+\end{center}
+
+\noindent\textbf{Motivation}
 
@@
 
-\noindent\textbf{\Large About the Data}
+\noindent\textbf{About the Data}
 
@@
 
-\newpage
-
-\noindent\textbf{\Large Exploratory Data Analysis}
+\noindent\textbf{Exploratory Data Analysis}
 
@@
 
-\noindent\textbf{\Large Full Model}
+\begin{center}
+\Large\textbf{\large Finding Our Model}
+\end{center}
+
+\noindent\textbf{Full Model}
 
@@
 
-\noindent\textbf{\Large Updated Model}
+\noindent\textbf{Updating Model: VIF and Lasso Considerations}
 
@@
 We saw that `gender`, `occupation`, `bmi_category`, and `sleep_disorder` converged to zero.
+Optimal lambda value was 0.0009622.
 
@@
       row.names = FALSE)
+
+# Create a tibble with the optimal lambda
+lambda_table <- tibble(
+  Description = "Optimal lambda (lambda.min)",
+  Value = cvmodelL$lambda.min
+)
+
+# Display in kable format
+kable(lambda_table, caption = "Optimal Lambda from Cross-Validation")
+
 ```
@@
 final_model <- lm(sleep_duration ~., sleep_test_final)
-final_model_summary <- summary(model_final)
+final_model_summary <- summary(final_model)
 
@@
 ```
+
+\begin{center}
+\Large\textbf{\large Discussion}
+\end{center}
+
+```{r include=FALSE}
+library(caret)
+library(knitr)
+library(dplyr)
+
+set.seed(510)
+
+# Set up 10-fold cross-validation
+train_control <- trainControl(method = "cv", number = 10)
+
+# 1. Initial model using full dataset
+cv_full <- train(sleep_duration ~ ., data = sleep, method = "lm", trControl = train_control)
+
+# 2. Final model using sleep_test_final dataset
+cv_final <- train(sleep_duration ~ ., data = sleep_test_final, method = "lm", trControl = train_control)
+
+# 3. Extract RMSE and R-squared
+cv_results <- tibble(
+  Model = c("Initial Model", "Final Model"),
+  RMSE = c(cv_full$results$RMSE, cv_final$results$RMSE),
+  Rsquared = c(cv_full$results$Rsquared, cv_final$results$Rsquared)
+)
+
+```
+
+```{r}
+# 4. Display as kable table
+kable(cv_results, caption = "10-Fold Cross-Validation Results for Initial and Final Models")
+```

### 2025-05-08 paper/paper.qmd
summary of full model
discussion

@@
 knitr::kable(final_model_stats, caption = "Final Model Statistics")
+
+final_model <- lm(sleep_duration ~ ., data = sleep_test_final)
+summary(final_model)
 ```
@@
 \end{center}
+
+Upon comparing the final model to the original full model, we observe a reduction in the Adjusted R² from 0.958 to 0.880, and an increase in the Residual Sum of Squares (RSS) from 9.28 to 27.61.
+The F-statistic increased slightly from 313.34 to 343.56, while the Total Sum of Squares remained nearly unchanged at approximately 236.
+This suggests that, although the simplified model explains slightly less of the variation in sleep duration, it still performs quite well in terms of overall fit and statistical significance.
+
+Despite the modest drop in Adjusted R², we are confident in selecting the final model.
+This model is much simpler, yet retains highly significant explanatory power, making it more interpretable and potentially more useful in real-world applications.
+The use of Lasso regularization, which penalizes more complex models, likely contributed to the decrease in Adjusted R².
+This trade-off was intentional.
+We prioritized a simpler model with a more strict variable selection.
+This has important practical benefits: a simpler model enables more efficient and cost-effective data collection for future studies.
+For example, if a healthcare provider wanted to screen individuals for sleep issues, collecting a smaller set of variables would be faster and less resource-demanding.
+In the final model, several predictors emerged as statistically significant at the 5% level, including: Physical Activity Level, Stress Levels 4 through 8, and Pulse Pressure (PP).
+
+These results align well with our expectations—individuals experiencing higher stress levels or reduced physical activity tend to sleep less.
+Interestingly, heart rate, while retained in the model, was not statistically significant at the 5% level, suggesting its role may be less direct in predicting sleepduration.
+Lastly, 10-fold cross-validation showed that both the full and final models performed well, with the final model achieving only a slight decrease in R² but with a meaningful gain in simplicity.
+This further supports our choice to recommend the final model as a strong, interpretable, and efficient tool for predicting sleep duration.
 

### 2025-05-08 paper/paper.qmd
adding in table for coefficients of final model

@@
 final_model <- lm(sleep_duration ~ ., data = sleep_test_final)
-summary(final_model)
+
+tidy(final_model) %>%
+  dplyr::select(term, estimate, p.value) %>%
+  kable(
+    digits = 3,
+    caption = "Coefficients and P-values from Linear Model",
+    booktabs = TRUE,
+    format = "latex",
+    escape = TRUE
+  )
+
 ```

### 2025-05-08 paper/paper.qmd
Update paper.qmd
added table and beta interpretation

@@
 ```
+
+```{r}
+library(knitr)
+library(tibble)
+
+# Create table manually
+final_model_table <- tribble(
+  ~Term, ~Estimate, ~`Standard Error`, ~Significance,
+  "(Intercept)", 14.8555, 0.6299, "***",
+  "physical_activity_level", 0.0158, 0.0015, "***",
+  "stress_level4", -1.3565, 0.0458, "***",
+  "stress_level5", -0.6208, 0.0491, "***",
+  "stress_level6", -0.9042, 0.0600, "***",
+  "stress_level7", -1.3989, 0.0627, "***",
+  "stress_level8", -1.6704, 0.0740, "***",
+  "heart_rate", -0.0401, 0.0059, "***",
+  "daily_steps", -0.0002, 0.0000, "***",
+  "PP", -0.0821, 0.0089, "***"
+)
+
+# Render as kable
+kable(
+  final_model_table,
+  caption = "Final Model Coefficients and Significance",
+  format = "pandoc"
+)
+```
+
+\begin{table}[ht]
+\centering
+\caption{Interpretation of Final Model Coefficients}
+\begin{tabular}{llp{8cm}}
+\toprule
+\textbf{Term} & \textbf{Estimate} & \textbf{Interpretation} \\
+\midrule
+Intercept & 14.86 & Baseline predicted sleep duration (in hours) when all predictors are at reference or zero levels. \\
+Physical Activity Level & 0.0158 & Each unit increase in physical activity is associated with an increase of approximately 0.016 hours (about 1 minute) of sleep. \\
+Stress Level 4 & -1.36 & Individuals with stress level 4 sleep 1.36 fewer hours than those at the reference level (e.g., stress level 1). \\
+Stress Level 5 & -0.62 & Stress level 5 corresponds to 0.62 fewer hours of sleep compared to the reference level. \\
+Stress Level 6 & -0.90 & Stress level 6 is associated with a reduction of 0.90 hours of sleep. \\
+Stress Level 7 & -1.40 & Individuals at stress level 7 sleep about 1.40 hours less than those at the reference level. \\
+Stress Level 8 & -1.67 & The highest stress level is linked to a reduction of 1.67 hours of sleep — the strongest negative effect. \\
+Heart Rate & -0.0401 & Each 1 bpm increase in heart rate is associated with 0.0401 fewer hours (~2.4 minutes) of sleep. \\
+Daily Steps & -0.0002 & Each additional step slightly reduces sleep. For example, 5,000 more steps corresponds to about 1 hour less sleep. \\
+Pulse Pressure (PP) & -0.0821 & Each 1 mmHg increase in pulse pressure corresponds to ~5 minutes less sleep. \\
+\bottomrule
+\end{tabular}
+\end{table}

### 2025-05-08 paper/paper.qmd
adding profile about best and worst sleep duration

@@
 Interestingly, heart rate, while retained in the model, was not statistically significant at the 5% level, suggesting its role may be less direct in predicting sleepduration.
+Based on the final model, individuals expected to have the longest sleep duration are those with high physical activity levels, low stress (level 1), lower heart rate, fewer daily steps, and lower pulse pressure (PP).
+In particular, physical activity shows a small but positive effect on sleep duration.
+Alternatively, the worst sleep duration is expected among individuals with very high stress levels (especially level 8), elevated heart rates, high PP, and excessive daily steps.
+These findings emphasize the importance of stress management and cardiovascular health in promoting better sleep.
 Lastly, 10-fold cross-validation showed that both the full and final models performed well, with the final model achieving only a slight decrease in R² but with a meaningful gain in simplicity.
